---
title: "Group Project Data"
output: html_document
date: "2023-02-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(here)
library(naniar)
library(shiny)
library(shinydashboard)
```

```{r}
getwd()
```

```{r}
diabetes_2019 <- read_csv("Group Project Data/diabetes_dataset__2019.csv") %>% clean_names()
diabetes <- read_csv("Group Project Data/diabetes.csv") %>% clean_names()
heart_failture <- read_csv("Group Project Data/heart_failure_clinical_records_dataset.csv") %>% clean_names()
```
## Check for NA
```{r}
any_na(diabetes_2019)
```

## NA overveiw
```{r}
naniar::miss_var_summary(diabetes_2019)
```

## Clean the data, make binary variables into numeric, countable variables. (Diabetes_2019)
```{r}
diabetes_2019_tidy <- diabetes_2019 %>% 
  mutate(high_bp_status = ifelse(high_bp == "yes", 1, 0)) %>% 
  mutate(family_heredity = ifelse(family_diabetes == "yes", 1, 0)) %>% 
  mutate(diabetic_status = ifelse(diabetic == "yes", 1, 0))
```

## Plots for high blood pressure: whether high blood pressure has relationship with diabetic. (Need more work)
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic_status != "NA") %>% 
  group_by(high_bp) %>%
  summarize(diabetic = sum(diabetic_status)) %>% 
  ggplot(aes(x=high_bp, y=diabetic, fill = high_bp))+
  geom_col(alpha = 0.6)+
  geom_text(aes(label=diabetic, vjust = -0.3), color = "deepskyblue4")+
  theme_linedraw()+
  labs(title = "High Blood Pressure vs. Diabetic", 
       x = "High Blood Pressure",
       y = "Diabetic")+
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic_status != "NA") %>% 
  count(diabetic_status, smoking) %>% 
  pivot_wider(names_from = diabetic_status,
              names_prefix = "diabetic_",
              values_from = n) %>% 
  mutate(diabetic_per = diabetic_1/(diabetic_1+diabetic_0)) %>% 
  ggplot(aes(x=smoking, y=diabetic_per, fill = smoking, label=sprintf("%0.2f", round(diabetic_per, digits = 2))))+
  geom_col(alpha = 0.6)+
  geom_text( vjust = -0.3, color = "deepskyblue4")+
  theme_linedraw()+
  labs(title = "Smoking vs. Diabetic", 
       x = "Smoking",
       y = "Diabetic Perentage")+
  theme(plot.title = element_text(hjust = 0.5))
```

## Relationship between physically active and diabetes: (Questions here, reorder, place them in order)
```{r}
diabetes_2019_tidy %>% 
  filter(physically_active != "NA") %>% 
  filter(diabetic_status != "NA") %>% 
  group_by(physically_active) %>% 
  summarize(diabetic = sum(diabetic_status)) %>% 
  ggplot(aes(x=physically_active, y=diabetic, fill = physically_active))+
  geom_col(alpha = 0.6)+
  scale_x_discrete(limits = c("none", "less than half an hr", "more than half an hr", "one hr or more"))+
  geom_text(aes(label=diabetic, vjust = -0.2), color = "deepskyblue4")+
  theme_linedraw()+
  labs(title = "Physically Active vs. Diabetic", 
       x = "Physically Active",
       y = "Diabetic")+
  theme(plot.title = element_text(hjust = 0.5))
```
## Bmi and Diabetes
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic_status!="NA") %>% 
  mutate(diabetic_status=as_factor(diabetic_status)) %>% 
  ggplot(aes(x=bmi, fill=diabetic_status))+
  geom_density(alpha=0.6)+
  theme_linedraw()+
  labs(title = "Bmi vs. Diabetic", 
       x = "Bmi",
       y = NULL)+
  theme(plot.title = element_text(hjust = 0.5))
```

# How bmi is connected to diabetes:
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic != "NA") %>% 
  ggplot(aes(x=diabetic, y=bmi, fill = diabetic))+
  geom_boxplot()+
  theme_linedraw()+
  scale_fill_discrete(labels=c("No", "Yes"))+
  labs(title = "Bmi vs. Diabetic", 
       x = NULL,
       y = NULL)+
  guides(fill = guide_legend(title = "Diabetes"))+
  theme(plot.title = element_text(hjust = 0.5))
```

## Mapping of diabetes percentage by counties
```{r}
library(usmap)
```

```{r}
here()
map_data_2004 <- read_csv(here("Tony Tao", "Group Project Data", "DiabetesAtlas_CountyData2004.csv"), skip = 2, na = "No Data")
map_data_2008 <- read_csv(here("Tony Tao", "Group Project Data", "DiabetesAtlas_CountyData2008.csv"), skip = 2, na = "No Data")
map_data_2012 <- read_csv(here("Tony Tao", "Group Project Data", "DiabetesAtlas_CountyData2012.csv"), skip = 2, na = "No Data")
map_data_2016 <- read_csv(here("Tony Tao", "Group Project Data", "DiabetesAtlas_CountyData2016.csv"), skip = 2, na = "No Data")
map_data_2020 <- read_csv(here("Tony Tao", "Group Project Data", "DiabetesAtlas_CountyData2020.csv"), skip = 2, na = "No Data")
```
## For 2004:
```{r}
glimpse(map_data_2004)
```

```{r}
map_data_2004$CountyFIPS <- as.numeric(map_data_2004$CountyFIPS)
```

```{r}
map_data_focused_2004 <- map_data_2004 %>% 
  select(CountyFIPS, Percentage) %>% 
  rename(fips = CountyFIPS, values = Percentage)
```

```{r}
summary(map_data_focused_2004$values)
```

```{r}
plot_usmap(data = map_data_focused_2004)+
  scale_fill_gradient2(low = "deepskyblue3", mid = "white", high = "red", name = "Diabetes Percentage", midpoint = 8)
```

## For 2008:
```{r}
map_data_2008$CountyFIPS <- as.numeric(map_data_2008$CountyFIPS)
```

```{r}
map_data_focused_2008 <- map_data_2008 %>% 
  select(CountyFIPS, Percentage) %>% 
  rename(fips = CountyFIPS, values = Percentage)
```

```{r}
summary(map_data_focused_2008$values)
```

```{r}
plot_usmap(data = map_data_focused_2004)+
  scale_fill_gradient2(low = "deepskyblue3", mid = "white", high = "red", name = "Diabetes Percentage", midpoint = 8)
```

## For 2012:
```{r}
map_data_2012$CountyFIPS <- as.numeric(map_data_2012$CountyFIPS)
```

```{r}
map_data_focused_2012 <- map_data_2012 %>% 
  select(CountyFIPS, Percentage) %>% 
  rename(fips = CountyFIPS, values = Percentage)
```

```{r}
summary(map_data_focused_2012$values)
```

```{r}
plot_usmap(data = map_data_focused_2012)+
  scale_fill_gradient2(low = "deepskyblue3", mid = "white", high = "red", name = "Diabetes Percentage", midpoint = 8)
```

## For 2016:
```{r}
map_data_2016$CountyFIPS <- as.numeric(map_data_2016$CountyFIPS)
```

```{r}
map_data_focused_2016 <- map_data_2016 %>% 
  select(CountyFIPS, Percentage) %>% 
  rename(fips = CountyFIPS, values = Percentage)
```

```{r}
summary(map_data_focused_2016$values)
```

```{r}
plot_usmap(data = map_data_focused_2016)+
  scale_fill_gradient2(low = "deepskyblue3", mid = "white", high = "red", name = "Diabetes Percentage", midpoint = 8)
```

## For 2020:
```{r}
map_data_2020$CountyFIPS <- as.numeric(map_data_2020$CountyFIPS)
```

```{r}
map_data_focused_2020 <- map_data_2020 %>% 
  select(CountyFIPS, Percentage) %>% 
  rename(fips = CountyFIPS, values = Percentage)
```

```{r}
summary(map_data_focused_2020$values)
```

```{r}
plot_usmap(data = map_data_focused_2020)+
  scale_fill_gradient2(low = "deepskyblue3", mid = "white", high = "red", name = "Diabetes Percentage", midpoint = 8)
```