---
title: "Group Project Data"
output: html_document
date: "2023-02-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(here)
library(naniar)
library(shiny)
library(shinydashboard)
```

```{r}
getwd()
```

```{r}
diabetes_2019 <- read_csv("Group Project Data/diabetes_dataset__2019.csv") %>% clean_names()
diabetes <- read_csv("Group Project Data/diabetes.csv") %>% clean_names()
heart_failture <- read_csv("Group Project Data/heart_failure_clinical_records_dataset.csv") %>% clean_names()
```
## Check for NA
```{r}
any_na(diabetes_2019)
```
## NA overveiw
```{r}
naniar::miss_var_summary(diabetes_2019)
```

## Clean the data, make binary variables into numeric, countable variables. (Diabetes_2019)
```{r}
diabetes_2019_tidy <- diabetes_2019 %>% 
  mutate(high_bp_status = ifelse(high_bp == "yes", 1, 0)) %>% 
  mutate(family_heredity = ifelse(family_diabetes == "yes", 1, 0)) %>% 
  mutate(diabetic_status = ifelse(diabetic == "yes", 1, 0))
```

## Plots for high blood pressure: whether high blood pressure has relationship with diabetic. 
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic_status != "NA") %>% 
  group_by(high_bp) %>%
  summarize(diabetic = sum(diabetic_status)) %>% 
  ggplot(aes(x=high_bp, y=diabetic, fill = high_bp))+
  geom_col(alpha = 0.6)+
  geom_text(aes(label=diabetic, vjust = -0.3), color = "deepskyblue4")+
  theme_linedraw()+
  labs(title = "High Blood Pressure vs. Diabetic", 
       x = "High Blood Pressure",
       y = "Diabetic")+
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
diabetes_2019_tidy %>% 
  count(diabetic_status, smoking)
```

```{r}
diabetes_2019_tidy %>% 
  filter(diabetic_status != "NA") %>% 
  group_by(smoking) %>%
  summarize(diabetic = sum(diabetic_status)) %>% 
  ggplot(aes(x=smoking, y=diabetic, fill = smoking))+
  geom_col(alpha = 0.6)+
  geom_text(aes(label=diabetic, vjust = -0.3), color = "deepskyblue4")+
  theme_linedraw()+
  labs(title = "High Blood Pressure vs. Diabetic", 
       x = "High Blood Pressure",
       y = "Diabetic")+
  theme(plot.title = element_text(hjust = 0.5))
```

## Relationship between physically active and diabetes: (Questions here, reorder, place them in order)
```{r}
diabetes_2019_tidy %>% 
  filter(physically_active != "NA") %>% 
  filter(diabetic_status != "NA") %>% 
  group_by(physically_active) %>% 
  summarize(diabetic = sum(diabetic_status)) %>% 
  ggplot(aes(x=physically_active, y=diabetic, fill = physically_active))+
  geom_col(alpha = 0.6)+
  scale_x_discrete(limits = c("none", "less than half an hr", "more than half an hr", "one hr or more"))+
  geom_text(aes(label=diabetic, vjust = -0.2), color = "deepskyblue4")+
  theme_linedraw()+
  labs(title = "Physically Active vs. Diabetic", 
       x = "Physically Active",
       y = "Diabetic")+
  theme(plot.title = element_text(hjust = 0.5))
```

## Relationship between bmi and diabetes:

# Distribution of bmi of the recorded people:
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic_status!="NA") %>% 
  mutate(diabetic_status=as_factor(diabetic_status)) %>% 
  ggplot(aes(x=bmi))+
  geom_density()+
  facet_wrap(~diabetic_status)
```

```{r}
diabetes_2019_tidy %>% 
  filter(diabetic_status!="NA") %>% 
  mutate(diabetic_status=as_factor(diabetic_status)) %>% 
  ggplot(aes(x=bmi, fill=diabetic_status))+
  geom_density(alpha=0.6)
```

# How bmi is connected to diabetes:
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic != "NA") %>% 
  ggplot(aes(x=diabetic, y=bmi))+
  geom_boxplot()
```

