---
title: "Group Project Data"
author: "Tony Tao"
output: 
  html_document: 
    keep_md: yes
date: "2023-02-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Library load
```{r, message = FALSE}
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(here)
library(naniar)
library(shiny)
library(shinydashboard)
```
# Get working directory
```{r}
getwd()
```
# Import Data
```{r}
diabetes_2019 <- read_csv("Group Project Data/diabetes_dataset__2019.csv") %>% clean_names()
```
Originally taken from Kaggle,
This dataset was collected by Neha Prerna Tigga and Dr. Shruti Garg of the Department of Computer Science and Engineering, BIT Mesra, Ranchi-835215 
Tigga, N. P., & Garg, S. (2020). Prediction of Type 2 Diabetes using Machine Learning Classification Methods. Procedia Computer Science, 167, 706-716. DOI: https://doi.org/10.1016/j.procs.2020.03.336

```{r}
heart_failture <- read_csv("Group Project Data/heart_failure_clinical_records_dataset.csv") %>% clean_names()
```
Originally taken from Kaggle,
Davide Chicco, Giuseppe Jurman: Machine learning can predict survival of patients with heart failure from serum creatinine and ejection fraction alone. BMC Medical Informatics and Decision Making 20, 16 (2020). https://doi.org/10.1186/s12911-020-1023-5

# Data overview
## Check for NA
```{r}
any_na(diabetes_2019)
```

```{r}
any_na(heart_failture)
```

## NA overveiw
```{r}
naniar::miss_var_summary(diabetes_2019)
```

```{r}
naniar::miss_var_summary(heart_failture)
```


## Clean the data(2019):
make binary variables into numeric, countable variables. (Diabetes_2019)
```{r}
diabetes_2019_tidy <- diabetes_2019 %>% 
  mutate(high_bp_status = ifelse(high_bp == "yes", 1, 0)) %>% 
  mutate(family_heredity = ifelse(family_diabetes == "yes", 1, 0)) %>% 
  mutate(diabetic_status = ifelse(diabetic == "yes", 1, 0))
```

## Clean the data(heart failture):
make binary variables into numeric, countable variables. (heart_failture)
```{r}
heart_failture_tidy <- heart_failture %>% 
  mutate(smoking_status = ifelse(smoking == 1, "yes", "no")) %>% 
  mutate(high_bp_status = ifelse(high_blood_pressure == 1, "yes", "no"))
```


# Data analysis and plotting:

## High bp vs. diabetes

###For diabetes_2019:
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic_status != "NA") %>% 
  count(diabetic_status, high_bp) %>% 
  pivot_wider(names_from = diabetic_status,
              names_prefix = "diabetic_",
              values_from = n) %>% 
  mutate(diabetic_per = diabetic_1/(diabetic_1+diabetic_0)) %>% 
  ggplot(aes(x=high_bp, y=diabetic_per, fill = high_bp, label=sprintf("%0.2f", round(diabetic_per, digits = 2))))+
  geom_col(alpha = 0.6)+
  scale_fill_discrete(labels=c("No", "Yes"))+
  geom_text(vjust = -0.3, color = "deepskyblue4")+
  theme_linedraw()+
  labs(title = "High bp vs. Diabetic", 
       x = NULL,
       y = "Diabetic Perentage")+
  guides(fill = guide_legend(title = "High bp"))+
  theme(plot.title = element_text(hjust = 0.5))
```
###For hear failture:
```{r}
heart_failture_tidy %>% 
  filter(diabetes != "NA") %>% 
  count(diabetes, high_bp_status) %>% 
  pivot_wider(names_from = diabetes,
              names_prefix = "diabetic_",
              values_from = n) %>% 
  mutate(diabetic_per = diabetic_1/(diabetic_1+diabetic_0)) %>% 
  ggplot(aes(x=high_bp_status, y=diabetic_per, fill = high_bp_status, label=sprintf("%0.2f", round(diabetic_per, digits = 2))))+
  geom_col(alpha = 0.6)+
  scale_fill_discrete(labels=c("No", "Yes"))+
  geom_text(vjust = -0.3, color = "deepskyblue4")+
  theme_linedraw()+
  labs(title = "High bp vs. Diabetic", 
       x = NULL,
       y = "Diabetic Perentage")+
  guides(fill = guide_legend(title = "High bp"))+
  theme(plot.title = element_text(hjust = 0.5))
```


## Smoking vs. diabetes

### For diabetes_2019:
Relationship between smoking and diabetic
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic_status != "NA") %>% 
  count(diabetic_status, smoking) %>% 
  pivot_wider(names_from = diabetic_status,
              names_prefix = "diabetic_",
              values_from = n) %>% 
  mutate(diabetic_per = diabetic_1/(diabetic_1+diabetic_0)) %>% 
  ggplot(aes(x=smoking, y=diabetic_per, fill = smoking, label=sprintf("%0.2f", round(diabetic_per, digits = 2))))+
  geom_col(alpha = 0.6)+
  scale_fill_discrete(labels=c("No", "Yes"))+
  geom_text(vjust = -0.3, color = "deepskyblue4")+
  theme_linedraw()+
  labs(title = "Smoking vs. Diabetic", 
       x = NULL,
       y = "Diabetic Perentage")+
  guides(fill = guide_legend(title = "Smoking"))+
  theme(plot.title = element_text(hjust = 0.5))
```
### For heart failture:
Relationship between smoking and diabetic
```{r}
heart_failture_tidy %>% 
  filter(diabetes != "NA") %>% 
  count(diabetes, smoking_status) %>% 
  pivot_wider(names_from = diabetes,
              names_prefix = "diabetic_",
              values_from = n) %>% 
  mutate(diabetic_per = diabetic_1/(diabetic_1+diabetic_0)) %>% 
  ggplot(aes(x=smoking_status, y=diabetic_per, fill = smoking_status, label=sprintf("%0.2f", round(diabetic_per, digits = 2))))+
  geom_col(alpha = 0.6)+
  scale_fill_discrete(labels=c("No", "Yes"))+
  geom_text(vjust = -0.3, color = "deepskyblue4")+
  theme_linedraw()+
  labs(title = "Smoking vs. Diabetic", 
       x = NULL,
       y = "Diabetic Perentage")+
  guides(fill = guide_legend(title = "Smoking"))+
  theme(plot.title = element_text(hjust = 0.5))
```





## Physically active vs. diabetes:
```{r}
diabetes_2019_tidy %>% 
  filter(physically_active != "NA") %>% 
  filter(diabetic_status != "NA") %>% 
  group_by(physically_active) %>% 
  summarize(diabetic = sum(diabetic_status)) %>% 
  ggplot(aes(x=physically_active, y=diabetic, fill = physically_active))+
  geom_col(alpha = 0.6)+
  scale_x_discrete(limits = c("none", "less than half an hr", "more than half an hr", "one hr or more"))+
  geom_text(aes(label=diabetic, vjust = -0.2), color = "deepskyblue4")+
  theme_linedraw()+
  labs(title = "Physically Active vs. Diabetic", 
       x = "Physically Active",
       y = "Diabetic")+
  theme(plot.title = element_text(hjust = 0.5))
```

## Bmi and diabetes
### Density plot
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic_status!="NA") %>% 
  mutate(diabetic_status=as_factor(diabetic_status)) %>% 
  ggplot(aes(x=bmi, fill=diabetic_status))+
  geom_density(alpha=0.6)+
  theme_linedraw()+
  scale_fill_discrete(labels=c("No", "Yes"))+
  labs(title = "Bmi vs. Diabetic", 
       x = "Bmi",
       y = NULL)+
  guides(fill = guide_legend(title = "Diabetes"))+
  theme(plot.title = element_text(hjust = 0.5))
```

### Box plot
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic != "NA") %>% 
  ggplot(aes(x=diabetic, y=bmi, fill = diabetic))+
  geom_boxplot(alpha = 0.6)+
  theme_linedraw()+
  scale_fill_discrete(labels=c("No", "Yes"))+
  labs(title = "Bmi vs. Diabetic", 
       x = NULL,
       y = NULL)+
  guides(fill = guide_legend(title = "Diabetes"))+
  theme(plot.title = element_text(hjust = 0.5))
```

## Mapping of diabetes percentage by counties
```{r}
library(usmap)
```
## Data import:
All the data is directly downloaded from the CDC United States Diabetes Surveillance System. 
```{r}
here()
map_data_2004 <- read_csv(here("Tony Tao", "Group Project Data", "DiabetesAtlas_CountyData2004.csv"), skip = 2, na = "No Data")
map_data_2008 <- read_csv(here("Tony Tao", "Group Project Data", "DiabetesAtlas_CountyData2008.csv"), skip = 2, na = "No Data")
map_data_2012 <- read_csv(here("Tony Tao", "Group Project Data", "DiabetesAtlas_CountyData2012.csv"), skip = 2, na = "No Data")
map_data_2016 <- read_csv(here("Tony Tao", "Group Project Data", "DiabetesAtlas_CountyData2016.csv"), skip = 2, na = "No Data")
map_data_2020 <- read_csv(here("Tony Tao", "Group Project Data", "DiabetesAtlas_CountyData2020.csv"), skip = 2, na = "No Data")
```
### For 2004:
```{r}
glimpse(map_data_2004)
```

```{r}
map_data_2004$CountyFIPS <- as.numeric(map_data_2004$CountyFIPS)
```

```{r}
map_data_focused_2004 <- map_data_2004 %>% 
  select(CountyFIPS, Percentage) %>% 
  rename(fips = CountyFIPS, values = Percentage)
```

```{r}
summary(map_data_focused_2004$values)
```

```{r}
plot_usmap(data = map_data_focused_2004)+
  scale_fill_gradient2(low = "deepskyblue3", mid = "white", high = "red", name = "Diabetes Percentage", midpoint = 8)
```

### For 2008:
```{r}
map_data_2008$CountyFIPS <- as.numeric(map_data_2008$CountyFIPS)
```

```{r}
map_data_focused_2008 <- map_data_2008 %>% 
  select(CountyFIPS, Percentage) %>% 
  rename(fips = CountyFIPS, values = Percentage)
```

```{r}
summary(map_data_focused_2008$values)
```

```{r}
plot_usmap(data = map_data_focused_2008)+
  scale_fill_gradient2(low = "deepskyblue3", mid = "white", high = "red", name = "Diabetes Percentage", midpoint = 8)
```

### For 2012:
```{r}
map_data_2012$CountyFIPS <- as.numeric(map_data_2012$CountyFIPS)
```

```{r}
map_data_focused_2012 <- map_data_2012 %>% 
  select(CountyFIPS, Percentage) %>% 
  rename(fips = CountyFIPS, values = Percentage)
```

```{r}
summary(map_data_focused_2012$values)
```

```{r}
plot_usmap(data = map_data_focused_2012)+
  scale_fill_gradient2(low = "deepskyblue3", mid = "white", high = "red", name = "Diabetes Percentage", midpoint = 8)
```

### For 2016:
```{r}
map_data_2016$CountyFIPS <- as.numeric(map_data_2016$CountyFIPS)
```

```{r}
map_data_focused_2016 <- map_data_2016 %>% 
  select(CountyFIPS, Percentage) %>% 
  rename(fips = CountyFIPS, values = Percentage)
```

```{r}
summary(map_data_focused_2016$values)
```

```{r}
plot_usmap(data = map_data_focused_2016)+
  scale_fill_gradient2(low = "deepskyblue3", mid = "white", high = "red", name = "Diabetes Percentage", midpoint = 8)
```

### For 2020:
```{r}
map_data_2020$CountyFIPS <- as.numeric(map_data_2020$CountyFIPS)
```

```{r}
map_data_focused_2020 <- map_data_2020 %>% 
  select(CountyFIPS, Percentage) %>% 
  rename(fips = CountyFIPS, values = Percentage)
```

```{r}
summary(map_data_focused_2020$values)
```

```{r}
plot_usmap(data = map_data_focused_2020)+
  scale_fill_gradient2(low = "deepskyblue3", mid = "white", high = "red", name = "Diabetes Percentage", midpoint = 8)
```