---
title: "Group Project Data"
output: html_document
date: "2023-02-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(here)
library(naniar)
library(shiny)
library(shinydashboard)
```

```{r}
getwd()
```

```{r}
diabetes_2019 <- read_csv("Group Project Data/diabetes_dataset__2019.csv") %>% clean_names()
diabetes <- read_csv("Group Project Data/diabetes.csv") %>% clean_names()
heart_failture <- read_csv("Group Project Data/heart_failure_clinical_records_dataset.csv") %>% clean_names()
```
## Check for NA
```{r}
any_na(diabetes_2019)
```
## NA overveiw
```{r}
naniar::miss_var_summary(diabetes_2019)
```



## Clean the data, make binary variables into numeric, countable variables. (Diabetes_2019)
```{r}
diabetes_2019_tidy <- diabetes_2019 %>% 
  mutate(high_bp_status = ifelse(high_bp == "yes", 1, 0)) %>% 
  mutate(family_heredity = ifelse(family_diabetes == "yes", 1, 0)) %>% 
  mutate(diabetic_status = ifelse(diabetic == "yes", 1, 0))
```

## Plots for high blood pressure: whether high blood pressure has relationship with diabetic. 
```{r}
diabetes_2019_tidy %>% 
  filter(diabetic_status != "NA") %>% 
  count(high_bp, diabetic_status) %>% 
  ggplot(aes(x=high_bp, y=n))+
  geom_col()
```

## Relationship between physically active and diabetes: (Questions here, reorder, place them in order)
```{r}
diabetes_2019_tidy %>% 
  filter(physically_active != "NA") %>% 
  count(physically_active, diabetic_status) %>% 
  ggplot(aes(x=reorder(physically_active, n), y=n))+
  geom_col()
```

## Relationship between bmi and diabetes:

# Distribution of bmi of the recorded people:
```{r}
diabetes_2019_tidy %>% 
  ggplot(aes(x=bmi))+
  geom_density()
```

# How bmi is connected to diabetes:
```{r}
diabetes_2019_tidy %>% 
  ggplot(aes(x=diabetic, y=bmi))+
  geom_boxplot()
```

